MultifidelityTrajectoryOptimization

### 2025-03-10 artefacts/paper.tex
write motivation and add experiments to extra section

The objective of this paper is to reduce the variance of the objective function estimation under sample average approximation.
A Monte Carlo(MC) simulation can be used to estimate the expected value of the objective function, but a control variate(CV) approach
would reduce the variance with the same computational cost. In the absence of an analytical expression of the expected value of the
objective function, an approximate control variate(ACV) approach is also proposed. The CV estimator will also serve as
the benchmark for the ACV estimator, since it is the best possible variance reduction that can be achieved with the ACV estimator.

The solution to the stochastic LQR controller, $U^*$ can be written as a function of the initial state $x_0$.

@@

Thus, we aim to reduce the variance of the objective function estimate thereby reducing the covariance of the
solution $U^*$ given a finite cost of computation.

@@

>>>>

For optimizing the trajectory of system under uncertainty, a common approach is to compute the expectation of the objective
function of a trajectory is by doing sample average approximation, where we propagate the trajectory with multiples samples
of the random variable, and take the average of all samples in a numerical optimizer to determine the optimal control inputs.
The most straighforward way to compute the expectation of the trajectory's objective function is to use Monte Carlo (MC) estimator.
When using a MC estimator, the variance of the estimated trajectory objective is related to the variance of the trajectory's
objective function in the following way.

$$
\text{Var}[S_n^{MC}(J)] = \frac{\text{Var}[J]}{n}
$$

Where $S_n^{MC}(J)$ is the MC estimator of the objective function $J$ with $n$ samples of the random variable.
The solution to the stochastic LQR controller, $U^*$ can be written as a function of the initial state $x_0$.

@@

The objective of this paper is to reduce the variance of the objective function estimation under sample average approximation,
thus, thereby reducing the covariance of the solution $U^*$ given a finite cost of computation.

A control variate (CV) estimator is a variance reduction technique that uses the correlation between the objective function
and another function to reduce the variance of the objective function estimation. Its variance has the form

$$
\text{Var}[S_{n_{cv}}^{CV}(J_h, J_l)] = \text{Var}[J_h] (1 - \rho^2)
$$

Where $J_h$ and $J_l$ are the high and low fidelity models of the objective function, $\rho$ is the Pearson correlation
coefficient between the high and low fidelity models, and $S_{n_{cv}}^{CV}(J_h, J_l)$ is the CV estimator of the objective
function $J_h$ using the low fidelity model $J_l$.

The CV estimator requires the analytical expectation of the expectation of the low fidelity model, and the covariance between
the high and low fidelity models. This might not be available in practice, and hence, an alternative is the approximate
control variate (ACV) estimator. The ACV estimator uses the statistical estimate of the expectation and covariance
of the low fidelity model, and has the variance

$$
\text{Var}[S_{n_{acv},m_{acv}}^{ACV}(J_h, J_l)] = \text{Var}[J_h] (1 - \frac{m_{acv}}{m_{acv} + n_{acv}} \rho^2)
$$

Where $m_{acv}$ and $n_{acv}$ are the number of samples used to estimate the expectation and covariance of the low fidelity model,

@@

\section{Extra}
To test the estimators, we set up a couple of experiments.

We fixed $\alpha = -1$ for the ACV estimator and increased the ratio $m:n$ for the ACV estimator
to see what the optimal sample allocation is
\begin{algorithm}
  \caption{ACV sample allocation experiment}
  \begin{algorithmic}
    \Require $J_h, J_l, x_0, e_s, ratio_{cost}, MC_{cost}, ratio_{mn}$
    \State $x_{0,samples} \gets \text{generate samples of } x_0 \text{ for all } e_s$
    \For{$r$ in $ratio_{mn}$}
    \State $n_{acv} \gets MC_{cost} / (1 + ratio_{cost} + r * ratio_{cost})$
    \State $m_{acv} \gets r * n_{acv}$
    \For{$x_0^{(i)}$ in $x_{0,samples}$}
    \State $O^{(i)} \gets S_{n_{acv},m_{acv}}^{ACV}(J_h, J_l, x_0^{(i)})$
    \EndFor
    \State $\text{Var}^{(r)} \gets \text{Var}[O]$
    \EndFor
  \end{algorithmic}
\end{algorithm}

$e_s$ is the number of estimator samples

$ratio_{cost}$ is the ratio of the cost of $J_l$ to $J_h$

$MC_{cost}$ is the equivalent cost of the MC estimator

Then we fixed $n$ for the ACV estimator and increased $m$ to see if it converges to the CV estimator variance
\begin{algorithm}
  \caption{ACV convergence experiment}
  \begin{algorithmic}
    \Require $J_h, J_l, x_0, e_s, ratio_{cost}, MC_{cost}, ms_{acv}$
    \State $x_{0,samples} \gets \text{generate samples of } x_0 \text{ for all } e_s$
    \State $n_{cv} \gets MC_{cost} / (1 + ratio_{cost})$
    \State $n_{acv} \gets MC_{cost} / (1 + ratio_{cost})$ (same as CV)
    \For{$m$ in $ms_{acv}$}
    \State $m_{acv} \gets m$
    \For{$x_0^{(i)}$ in $x_{0,samples}$}
    \State $O_{cv}^{(i)} \gets S_{n_{cv}}^{CV}(J_h, J_l, x_0^{(i)})$
    \State $O_{acv}^{(i)} \gets S_{n_{acv},m_{acv}}^{ACV}(J_h, J_l, x_0^{(i)})$
    \EndFor
    \State $\text{Var}_{cv}^{(m)} \gets \text{Var}[O_{cv}]$
    \State $\text{Var}_{acv}^{(m)} \gets \text{Var}[O_{acv}]$
    \State $\text{Var}^{(m)} \gets (\text{Var}_{acv}^{(m)} - \text{Var}_{cv}^{(m)})^2$
    \EndFor
  \end{algorithmic}
\end{algorithm}
